Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
capitalization: BOF -> BoF

Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics. In this context, a classical approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.

It is worth mentionning that this task typically falls into a unsupervised paradigm where no prior knowldge is used to model a given scene. For each scene $s$, a model is compute $M_s$ is computed. The similarity among the scene $s_1$ and $s_2$ is computed as the similarity, see Section~\ref{sec:experiments} for further details. The BOF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}. In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set. In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed. The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than the other likelihoods.

While BOF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
The latter approach should therefore prove more fruitful in measuring auditory scene similarity.

@@

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF models which treat each observation separately and do not capture their temporal structure. A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.

@@

The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The distance on this average feature vector is then used to determine $p@k$. For the BOF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}. The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation. To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset. the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$. Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features. Recommended number of Gaussians for MFCCs given in~\cite{aucouturier2007bag} is 10.

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BOF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).

@@

Results for the acoustic scene similarity retrieval task demonstrate that logarithmically compressed scattering features outerform MFCCs.
Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BOF and summary statistic measures.

@@

As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BOF and early integration schemes.
This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information~\cite{Anden2014}.
We also note that applying PCA reduction to the scattering transform has little effect for the early integration scheme.
In the context of summary statistics, $30$ dimensions are sufficient to discriminate between auditory scenes.

Comparing the BOF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
This is in line with previous results on BOF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.

@@

For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes. However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids. This result is in line with the findings of~\cite{lagrange:hal-01082501}. To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects. The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BOF model, as seen previously.

@@

This paper presents a new approach for modeling acoustic scenes based on scattering transforms at small scales and cluster-based representations at large scales. Compared to traditional BOF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. To adequately capture such distinct events, we develop a cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and emerging tasks in ecoacoustics~\cite{sueur2015ecoacoustics, wimmer2013sampling}.

>>>>

Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics. In this context, a classical approach is the bag of frames (BoF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.

It is worth mentionning that this task typically falls into a unsupervised paradigm where no prior knowldge is used to model a given scene. For each scene $s$, a model is compute $M_s$ is computed. The similarity among the scene $s_1$ and $s_2$ is computed as the similarity, see Section~\ref{sec:experiments} for further details. The BoF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}. In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set. In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed. The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than the other likelihoods.

While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
The latter approach should therefore prove more fruitful in measuring auditory scene similarity.

@@

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BoF models which treat each observation separately and do not capture their temporal structure. A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.

@@

The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The distance on this average feature vector is then used to determine $p@k$. For the BoF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}. The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation. To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset. the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$. Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features. Recommended number of Gaussians for MFCCs given in~\cite{aucouturier2007bag} is 10.

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).

@@

Results for the acoustic scene similarity retrieval task demonstrate that logarithmically compressed scattering features outerform MFCCs.
Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BoF and summary statistic measures.

@@

As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BoF and early integration schemes.
This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information~\cite{Anden2014}.
We also note that applying PCA reduction to the scattering transform has little effect for the early integration scheme.
In the context of summary statistics, $30$ dimensions are sufficient to discriminate between auditory scenes.

Comparing the BoF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
This is in line with previous results on BoF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.

@@

For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes. However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids. This result is in line with the findings of~\cite{lagrange:hal-01082501}. To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects. The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BoF model, as seen previously.

@@

This paper presents a new approach for modeling acoustic scenes based on scattering transforms at small scales and cluster-based representations at large scales. Compared to traditional BoF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. To adequately capture such distinct events, we develop a cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BoF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and emerging tasks in ecoacoustics~\cite{sueur2015ecoacoustics, wimmer2013sampling}.
